### 2018-07-05 paper/5-implementations.tex
Clarify topological implementation

\begin{figure}
\begin{minted}[fontsize=\small]{haskell}
-- Make build system; stores current time and file modification times
type Time       = Integer
type MakeInfo k = (Time, Map k Time)
\end{minted}
\vspace{0mm}
\begin{minted}[fontsize=\small]{haskell}
make :: Ord k => Build Applicative (MakeInfo k) k v
make = topological modTimeRebuilder
\end{minted}
\vspace{0mm}
\begin{minted}[fontsize=\small]{haskell}
-- A task rebuilder based on file modification times
modTimeRebuilder :: Ord k => Rebuilder Applicative (MakeInfo k) k v
modTimeRebuilder key value task = Task $ \fetch -> do
    (now, modTimes) <- get
    let dirty = case Map.lookup key modTimes of
            Nothing -> True
            time -> any (\d -> Map.lookup d modTimes > time) (dependencies task)
    if not dirty then return value else do
        put (now + 1, Map.insert key now modTimes)
        run task fetch
\end{minted}
\vspace{0mm}
\begin{minted}[fontsize=\small]{haskell}
-- A topological task scheduler
topological :: Ord k => Scheduler Applicative i i k v
topological rebuilder tasks target = execState $ forM_ order $ \key -> case tasks key of
    Nothing   -> return ()
    Just task -> do
        store <- get
        let newTask :: Task (MonadState i) k v
            newTask = rebuilder key (getValue key store) task
            fetch :: k -> State i v
            fetch k = return (getValue k store)
        newValue <- liftStore (run newTask fetch)
        modify $ putValue key newValue
  where
    deps k = case tasks k of { Nothing -> [@@]; Just task -> dependencies task }
    order  = fromJust $ topSort (reachable deps target) -- Assumes acyclic dependencies
\end{minted}
\vspace{0mm}
\begin{minted}[fontsize=\small]{haskell}
-- Standard graph algorithms (implementation omitted)
reachable :: Ord k => (k -> [k]) -> k -> Graph k
topSort   :: Ord k => Graph k -> Maybe [k] -- Returns Nothing on a cyclic graph
\end{minted}
\vspace{0mm}
\begin{minted}[fontsize=\small]{haskell}
-- Expand the scope of visibility of a stateful computation
liftStore :: State i a -> State (Store i k v) a
liftStore x = do (a, newInfo) <- gets (runState x . getInfo)
                 modify (putInfo newInfo)
                 return a
\end{minted}
\vspace{-3mm}
\caption{An implementation of \Make using our framework.}\label{fig-make-implementation}
\vspace{-3mm}
\end{figure}

@@

\simon{topological is hard to read because of the execState and forM stuff.
  Could we have a local defn, with a type signature, for the $\lambda key. blah$?
  If necessary we could omit the code for liftStore, leaving only its signature}
\Make's scheduler, \hs{topological}, processes keys in a linear \hs{order} based on a
topological sort of the statically known dependency graph
(see~\S\ref{sec-parallelism} for parallel \Make). Our definition in
Fig.~\ref{fig-make-implementation} is polymorphic with respect to the type of
build information \hs{i} and is therefore compatible with any applicative
\hs{rebuilder}. The scheduler calls the supplied \hs{rebuilder} on every
\hs{key} in the \hs{order}, and runs the obtained \hs{newTask} to compute the
\hs{newValue}. Note that \hs{newTask} has access only to the \hs{i} part of the
\hs{Store}~\hs{i}~\hs{k}~\hs{v}, but the rest of the \hs{do} block runs in the
\hs{State}~\hs{(@@Store}~\hs{i}~\hs{k}~\hs{v)} monad; we use the (unremarkable)
helper function \hs{liftStore} to fix the mismatch. The \hs{newTask} finds
values of the \hs{key}'s dependencies via the \hs{fetch} callback, which is
defined to directly read the \hs{store}.

The \hs{where} clause corresponds to the pre-processing stage, which uses the
function \hs{dependencies}, defined in~\S\ref{sec-deps}, to extract static
dependencies from the provided applicative \hs{task}. We compute the linear
processing \hs{order} by constructing the graph of keys \hs{reachable} from the
\hs{target} via dependencies, and performing the topological sort of the result.
We omit implementation of textbook graph algorithms
\hs{reachable} and \hs{topSort}, e.g. see~\cite{cormen2001introduction}.

>>>>

\begin{figure}
\begin{minted}[fontsize=\small]{haskell}
-- Make build system; stores current time and file modification times
type Time       = Integer
type MakeInfo k = (Time, Map k Time)
\end{minted}
\vspace{0mm}
\begin{minted}[fontsize=\small]{haskell}
make :: Ord k => Build Applicative (MakeInfo k) k v
make = topological modTimeRebuilder
\end{minted}
\vspace{0mm}
\begin{minted}[fontsize=\small]{haskell}
-- A task rebuilder based on file modification times
modTimeRebuilder :: Ord k => Rebuilder Applicative (MakeInfo k) k v
modTimeRebuilder key value task = Task $ \fetch -> do
    (now, modTimes) <- get
    let dirty = case Map.lookup key modTimes of
            Nothing -> True
            time -> any (\d -> Map.lookup d modTimes > time) (dependencies task)
    if not dirty then return value else do
        put (now + 1, Map.insert key now modTimes)
        run task fetch
\end{minted}
\vspace{0mm}
\begin{minted}[fontsize=\small]{haskell}
-- A topological task scheduler
topological :: Ord k => Scheduler Applicative i i k v
topological rebuilder tasks target = execState $ mapM_ build order
  where
    build :: k -> State (Store i k v) ()
    build key = case tasks key of
        Nothing -> return ()
        Just task -> do
            store <- get
            let value = getValue key store
                newTask :: Task (MonadState i) k v
                newTask = rebuilder key value task
                fetch :: k -> State i v
                fetch k = return (getValue k store)
            newValue <- liftStore (run newTask fetch)
            modify $ updateValue key value newValue
    order = case topSort (graph deps target) of
        Nothing -> error "Cannot build tasks with cyclic dependencies"
        Just xs -> xs
    deps k = case tasks k of { Nothing -> []; Just task -> dependencies task }
\end{minted}
\vspace{0mm}
\begin{minted}[fontsize=\small]{haskell}
-- Standard graph algorithms (implementation omitted)
reachable :: Ord k => (k -> [k]) -> k -> Graph k
topSort   :: Ord k => Graph k -> Maybe [k] -- Returns Nothing on a cyclic graph
\end{minted}
\vspace{0mm}
\begin{minted}[fontsize=\small]{haskell}
-- Expand the scope of visibility of a stateful computation
liftStore :: State i a -> State (Store i k v) a
liftStore x = do (a, newInfo) <- gets (runState x . getInfo)
                 modify (putInfo newInfo)
                 return a
\end{minted}
\vspace{-3mm}
\caption{An implementation of \Make using our framework.}\label{fig-make-implementation}
\vspace{-3mm}
\end{figure}

@@

\Make's scheduler, \hs{topological}, processes keys in a linear \hs{order} based
on a topological sort of the statically known dependency graph
(see~\S\ref{sec-parallelism} for parallel \Make). Our definition in
Fig.~\ref{fig-make-implementation} is polymorphic with respect to the type of
build information \hs{i} and is therefore compatible with any applicative
\hs{rebuilder}. The scheduler calls the supplied \hs{rebuilder} on every
\hs{key} in the \hs{order}, and runs the obtained \hs{newTask} to compute the
\hs{newValue}. Note that \hs{newTask} has access only to the \hs{i} part of the
\hs{Store}~\hs{i}~\hs{k}~\hs{v}, but the rest of the \hs{do} block runs in the
\hs{State}~\hs{(@@Store}~\hs{i}~\hs{k}~\hs{v)} monad; we use the (unremarkable)
helper function \hs{liftStore} to fix the mismatch. The \hs{newTask} finds
values of the \hs{key}'s dependencies via the \hs{fetch} callback, which is
defined to directly read the \hs{store}.

The pre-processing stage uses the function \hs{dependencies}, defined
in~\S\ref{sec-deps}, to extract static dependencies from the provided
applicative \hs{task}. We compute the linear processing \hs{order} by
constructing the graph of keys \hs{reachable} from the \hs{target} via
dependencies, and performing the topological sort of the result. We omit
implementation of textbook graph algorithms \hs{reachable} and \hs{topSort},
e.g. see~\cite{cormen2001introduction}.
